---
title: "Play_Selection"
author: "mdgallery"
date: "6/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Play_Selection

# Load necessary and useful packages. 

```{r}
library(tidyverse)
library(nflfastR)
library(ggrepel)
library(ggimage)
library(gsisdecoder)
library(sqldf)

options(scipen = 9999)
```


# Generate data needed - full game data for 2020, weekly player stats, season player stats, and roster data.

```{r}
games_2020 <- nflfastR::load_pbp(2020) %>%
  dplyr::filter(season_type == "REG", !is.na(down), !(play_type == "no_play"))

player_stats_weekly_2020 <- games_2020 %>%
  nflfastR::calculate_player_stats(weekly = TRUE)

player_stats_season_2020 <- games_2020 %>%
  nflfastR::calculate_player_stats(weekly = FALSE)
  
roster_2020 <- nflfastR::fast_scraper_roster(2020) %>%
  select(season, team, position, depth_chart_position, full_name, first_name, last_name, birth_date, gsis_id, yahoo_id, headshot_url)
```

# Create full tables of individual player stats, both by season and by individual week performances, that includes roster information.

```{r}
full_weekly_2020 <- roster_2020 %>%
  dplyr::full_join(player_stats_weekly_2020, by = c("gsis_id" = "player_id"))

full_season_2020 <- roster_2020 %>%
  dplyr::full_join(player_stats_season_2020, by = c("gsis_id" = "player_id"))
```

# Load team and position information into the big dataframe.

```{r}
select_games_2020 <- games_2020 %>%
  dplyr::select(game_id, home_team, away_team, season_type, week, stadium, weather, roof, surface,
                play_id, yardline_100, game_seconds_remaining, drive, qtr, down, ydstogo, 
                play_type, play_type_nfl, posteam, posteam_type, defteam, first_down, 
                td_team, total_home_score, total_away_score, posteam_score, defteam_score, score_differential,
                desc,
                pass_attempt, complete_pass, incomplete_pass, pass_touchdown, interception, passing_yards, 
                passer, passer_id,
                fumble_lost, fumbled_1_player_id, fumbled_1_player_name,
                receiving_yards, receiver, receiver_id, 
                rush_attempt, run_location, run_gap, rushing_yards, rush_touchdown, rusher, rusher_id)

roster_short_2020 <- roster_2020 %>%
  dplyr::select(team, position, depth_chart_position, gsis_id)

new1 <- select_games_2020 %>%
  left_join(roster_short_2020, by = c("passer_id" = "gsis_id"), na_matches = "never") %>%
  dplyr::rename(passer_team = team, passer_position = position, passer_depth = depth_chart_position)

new2 <- new1 %>%
  left_join(roster_short_2020, by = c("receiver_id" = "gsis_id"), na_matches = "never") %>%
  dplyr::rename(receiver_team = team, receiver_position = position, receiver_depth = depth_chart_position)

new3 <- new2 %>%
  left_join(roster_short_2020, by = c("rusher_id" = "gsis_id"), na_matches = "never") %>%
  dplyr::rename(rusher_team = team, rusher_position = position, rusher_depth = depth_chart_position)

# later dumped the depth_chart_positions columns due to volume of NA's

full_weekly_with_positions_2020 <- new3 %>%
  dplyr::select(game_id, home_team, away_team, season_type, week, stadium, weather, roof, surface,
                play_id, yardline_100, game_seconds_remaining, drive, qtr, down, ydstogo, 
                play_type, play_type_nfl, posteam, posteam_type, defteam, first_down, 
                td_team, total_home_score, total_away_score, posteam_score, defteam_score, score_differential,
                desc,
                pass_attempt, complete_pass, incomplete_pass, pass_touchdown, interception, passing_yards, 
                passer, passer_id, passer_team, passer_position,
                fumble_lost, fumbled_1_player_id, fumbled_1_player_name,
                receiving_yards, receiver, receiver_id, receiver_team, receiver_position,
                rush_attempt, run_location, run_gap, rushing_yards, rush_touchdown, rusher, rusher_id,
                rusher_team, rusher_position)
```


```{r}
data <- full_weekly_with_positions_2020

play_choice <- data %>%
  dplyr::filter(play_type == "run" | play_type == "pass") %>%
  dplyr::group_by(down, ydstogo, yardline_100, score_differential)
play_choice
```

```{r}
play_choice <- data %>%
  dplyr::filter(play_type == "run" | play_type == "pass") %>%
  dplyr::count(posteam, down, ydstogo, yardline_100, score_differential, play_type) %>%
  dplyr::mutate(bin_yardline_100 = cut_interval(yardline_100, length = 10)) %>%
  dplyr::mutate(bin_score_differential = cut_interval(score_differential, length = 8, center = 0)) %>%
  dplyr::mutate(nplays = as.numeric(n)) %>%
  dplyr::filter(ydstogo < 11)
  #dplyr::select(down, ydstogo, bin_yardline_100, bin_score_differential, play_type, n)
#play_choice

run_play <- data %>%
  dplyr::filter(play_type == "run") %>%
  dplyr::count(posteam, down, ydstogo, yardline_100, score_differential, rusher) %>%
  dplyr::mutate(bin_yardline_100 = cut_interval(yardline_100, length = 10)) %>%
  dplyr::mutate(bin_score_differential = cut_interval(score_differential, length = 8, center = 0)) %>%
  dplyr::mutate(nplays = as.numeric(n)) %>%
  dplyr::filter(ydstogo < 11)

pass_play <- data %>%
  dplyr::filter(play_type == "pass") %>%
  dplyr::count(posteam, down, ydstogo, yardline_100, score_differential, passer, receiver) %>%
  dplyr::mutate(bin_yardline_100 = cut_interval(yardline_100, length = 10)) %>%
  dplyr::mutate(bin_score_differential = cut_interval(score_differential, length = 8, center = 0)) %>%
  dplyr::mutate(nplays = as.numeric(n)) %>%
  dplyr::filter(ydstogo < 11)

```

## Gain an idea on trends of play selection based on the down, yards to go, and field position

```{r}
g <- ggplot(data = play_choice)
g + geom_jitter(aes(x = ydstogo, y = yardline_100, size = nplays, color = play_type)) +
  scale_size_area("Number of plays", breaks = c(50,100))
  facet_wrap(~ down)
```

## Let's check for vultures on each team

First we need to create separate dataframes for each team

```{r}
#teams <- c("SF", "ARI", "CHI", "DET", "CLE", "BAL", "LA",  "DAL", "GB",  "MIN", "HOU", "KC",  "IND", "JAX", "CIN", "LAC", "LV",  "CAR", "MIA", "NE",  "NYJ", "BUF", "WAS", "PHI", "NYG", "PIT", "ATL", "SEA", "NO", "TB", "TEN", "DEN")

teams <- c(as.list(unique(full_weekly_with_positions_2020[c("posteam")])))
#teams

# Test case
mia_play_choice <- data %>%
  dplyr::filter(posteam == "MIA") %>%
  dplyr::filter(play_type == "run" | play_type == "pass") %>%
  dplyr::count(posteam, down, ydstogo, yardline_100, score_differential, play_type) %>%
  dplyr::mutate(bin_yardline_100 = cut_interval(yardline_100, length = 10)) %>%
  dplyr::mutate(bin_score_differential = cut_interval(score_differential, length = 8, center = 0)) %>%
  dplyr::mutate(nplays = as.numeric(n)) %>%
  dplyr::filter(ydstogo < 11)

# Find rushers with >10 carries in the season
mia_rushers_10_rushes <- data %>%
  dplyr::filter(play_type == "run" & posteam == "MIA") %>%
  dplyr::group_by(rusher) %>%
  dplyr::summarise(rushes = n()) %>%
  dplyr::ungroup() %>%
  dplyr::filter(rushes > 10)
mr10r <- pull(mia_rushers_10_rushes, rusher)
mia_rushers_10_rushes

mia_run_play <- data %>%
  dplyr::filter(play_type == "run" & posteam == "MIA" & rusher %in% mr10r) %>%
#  dplyr::group_by(posteam, down, ydstogo, yardline_100, score_differential, rusher) %>%
#  dplyr::ungroup() %>%
#  dplyr::count(rusher, down, ydstogo, yardline_100, score_differential) %>%
  dplyr::mutate(bin_yardline_100 = cut_interval(yardline_100, length = 10)) %>%
  dplyr::mutate(bin_score_differential = cut_interval(score_differential, length = 8, center = 0)) %>%
#  dplyr::mutate(nplays = as.numeric(n)) %>%
  dplyr::filter(ydstogo < 11)
mia_run_play$rusher <- ifelse(is.na(mia_run_play$rusher), mia_run_play$passer, mia_run_play$rusher)
mia_run_play$rusher_position <- ifelse(is.na(mia_run_play$rusher_position), mia_run_play$passer_position, mia_run_play$rusher_position)
mia_run_play

#mia_play_choice

g <- ggplot(data = mia_run_play)
g + geom_jitter(aes(x = ydstogo, y = yardline_100, color = rusher)) +
  scale_size_area("Number of plays", breaks = c(25,50)) +
  facet_grid(~ down, scales = "free_x")
```

#####
team_play_choice <- function(x) {
  for (i in teams) {
    assign(paste(teams[i]), "_play_choice")
  }
  mia_play_choice <- data %>%
  dplyr::filter(posteam == "MIA") %>%
  dplyr::filter(play_type == "run" | play_type == "pass") %>%
  dplyr::count(posteam, down, ydstogo, yardline_100, score_differential, play_type) %>%
  dplyr::mutate(bin_yardline_100 = cut_interval(yardline_100, length = 10)) %>%
  dplyr::mutate(bin_score_differential = cut_interval(score_differential, length = 8, center = 0)) %>%
  dplyr::mutate(nplays = as.numeric(n)) %>%
  dplyr::filter(ydstogo < 11)
}
#####




```{r, eval=FALSE}
g <- ggplot(data = play_choice,aes(x = ydstogo, y = yardline_100, color = play_type, size = nplays))

#g + geom_jitter(aes(x = ydstogo, y = yardline_100, size = nplays, color = play_type)) +
#  facet_wrap(~ down)

#g + stat_density2d(geom="tile", aes(fill=..density..), contour = FALSE) +
#  geom_point(color="white")

#g + geom_violin(aes(group = ydstogo)) + 
#  geom_jitter()

g + geom_violin(aes(group = down, alpha = ydstogo), adjust = .2) +
  facet_grid(play_type ~ down)

g + geom_violin(aes(group = down, alpha = ydstogo), adjust = .2, scale = "count") +
  facet_grid(play_type ~ down)

g + geom_violin(aes(group = down, alpha = ydstogo, fill = play_type), adjust = .2) +
  facet_grid(play_type ~ down)
```

```{r, eval=FALSE}
g <- ggplot(data = play_choice,aes(x = ydstogo, y = yardline_100, color = play_type, size = nplays))

g + geom_jitter(shape = 1) +
  scale_size_area("Number of plays", breaks = c(50,100)) +
  facet_wrap(~ ydstogo, scales = "free_x")

h <- ggplot(data = play_choice,aes(x = down, y = yardline_100, color = play_type, size = nplays))

h + geom_jitter(shape = 1) +
  scale_size_area("Number of plays", breaks = c(50,100)) +
  facet_wrap(~ ydstogo)
```

